### 2017-04-11 draft/draft.tex
abstract update

% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
\documentclass{llncs}
%
\usepackage{makeidx}  % allows for indexgeneration
% For figures
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subfigure} 
\usepackage{color}
%\usepackage{subfig}%[caption=false,font=footnotesize]

% For citations
\usepackage{natbib}

% For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}

% As of 2011, we use the hyperref package to produce hyperlinks in the
% resulting PDF.  If this breaks your system, please commend out the
% following usepackage line and replace \usepackage{icml2012} with
% \usepackage[nohyperref]{icml2012} above.
\usepackage{hyperref}

% Packages hyperref and algorithmic misbehave sometimes.  We can fix
% this with the following command.
\newcommand{\theHalgorithm}{\arabic{algorithm}}

%-------------------------------------------------------------
%                      Own Commands
%-------------------------------------------------------------
\usepackage{amsmath}
\newcommand{\cbn}{\textsc{Cbn}}
\newcommand{\bn}{\textsc{Bn}}
\newcommand{\huang}[1]{\textcolor{blue}{#1}}
\newcommand{\mycomment}[1]{\textcolor{red}{( ...#1 )}}
\renewcommand{\algorithmiccomment}[1]{/* #1 */}
\def\ci{\perp\!\!\!\perp}
\def\dep{\perp\!\!\!\perp\!\!\!\!\!\!\!/\,\,\,\,}
% Theorem & Co environments and counters
%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{lemma}[theorem]{Lemma}
%\newtheorem{corollary}[theorem]{Corollary}
%\newtheorem{remark}[theorem]{Remark}
%\newtheorem{definition}[theorem]{Definition}
\newtheorem{equat}[theorem]{Equation}
%\newtheorem{example}[theorem]{Example}
%-------------------------------------------------------------

\begin{document}
%
\frontmatter          % for the preliminaries
%
%\pagestyle{headings}  % switches on printing of running heads
%\addtocmark{Hamiltonian Mechanics} % additional mark in the TOC
%%
%\tableofcontents
%
\mainmatter              % start of the contributions
%
\title{ Communication efficient learning from Multiple Experts}
%
%\titlerunning{Hamiltonian Mechanics}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Bojan  Kolosnjaji\inst{1} \and Huang Xiao\inst{2} 
	\and Claudia Eckert\inst{1}\inst{2} }
%Jeffrey Dean \and David Grove \and Craig Chambers \and Kim~B.~Bruce \and
%Elsa Bertino}
%
\authorrunning{Bojan Kolosnjaji et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
%\tocauthor{Ivar Ekeland, Roger Temam, Jeffrey Dean, David Grove,
%Craig Chambers, Kim B. Bruce, and Elisa Bertino}
%
\institute{Technical University of Munich, Fraunhofer AISEC\\
\email{kolosnjaji@sec.in.tum.de, huang.xiao@aisec.fraunhofer.de, eckert@sec.in.tum.de}
%\\ WWW home page:
%\texttt{https://www.aisec.fraunhofer.de/}
%\and
%Universit\'{e} de Paris-Sud,
%Laboratoire d'Analyse Num\'{e}rique, B\^{a}timent 425,\\
%F-91405 Orsay Cedex, France
}

\maketitle              % typeset the title of the contribution

\begin{abstract}
Learning from multiple sources of labeled data can increase the confidence of a machine learning system, 
if the expertise of different labelers can be assembled efficiently. However,
using a large amount of labelers in the test time can be expensive in the long run. For instance, in crowdsourcing we have to pay each labeler in order to get the annotations. In % a scenario of 
Internet-of-Things, 
obtaining data from multiple sensors for direct fusion 
incurs high communication overhead. 
%Nevertheless, in many situations a small subset of 
%%competent experts 
%data sources could already 
%%constitute a very accurate labeling system
%deliver reliable annotations,

%Therefore we need to reduce the cost of communication with other labelers effectively}. 
In this paper, we 
%design a methodolojgy
propose a method for training a classification system 
%where data labels are obtained from a small set of labelers and the labelers are weighted optimally in order to offer high classification accuracy. 
from data labeled by multiple annotators, where only a small subset of annotators are optimally chosen to reduce communication effort. We define an iterative optimization procedure where the sparsity of a client weight vector is enforced, 
%is controlled 
while reducing the overall classification error. 
%Using this procedure, 
Using a set of experiments, we show that our approach 
%obtain results close to 
can pertain the classification performance as on the complete set of labelers, while reducing 
%the number of those 
the communication effort by over 60\% on several different tasks.

%submit  abstract until 13.04

\keywords{machine learning, communication-efficiency, crowdsourcing, Internet-of-Things}
\end{abstract}
%

% ---- Bibliography ----
%
\bibliographystyle{plain}
\bibliography{literature} % use this
%\clearpage
%\addtocmark[2]{Author Index} % additional numbered TOC entry
%\renewcommand{\indexname}{Author Index}
%\printindex
%\clearpage
%\addtocmark[2]{Subject Index} % additional numbered TOC entry
%\markboth{Subject Index}{Subject Index}
%\renewcommand{\indexname}{Subject Index}
%\input{subjidx.ind}
\end{document}