### 2017-04-12 draft/draft.tex
draft introduction

% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
\documentclass{llncs}
%
\usepackage{makeidx}  % allows for indexgeneration
% For figures
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subfigure} 
\usepackage{color}
%\usepackage{subfig}%[caption=false,font=footnotesize]

% For citations
\usepackage{natbib}

% For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}

% As of 2011, we use the hyperref package to produce hyperlinks in the
% resulting PDF.  If this breaks your system, please commend out the
% following usepackage line and replace \usepackage{icml2012} with
% \usepackage[nohyperref]{icml2012} above.
\usepackage{hyperref}

% Packages hyperref and algorithmic misbehave sometimes.  We can fix
% this with the following command.
\newcommand{\theHalgorithm}{\arabic{algorithm}}

%-------------------------------------------------------------
%                      Own Commands
%-------------------------------------------------------------
\usepackage{amsmath}
\newcommand{\cbn}{\textsc{Cbn}}
\newcommand{\bn}{\textsc{Bn}}
\newcommand{\huang}[1]{\textcolor{blue}{#1}}
\newcommand{\mycomment}[1]{\textcolor{red}{( ...#1 )}}
\renewcommand{\algorithmiccomment}[1]{/* #1 */}
\def\ci{\perp\!\!\!\perp}
\def\dep{\perp\!\!\!\perp\!\!\!\!\!\!\!/\,\,\,\,}
% Theorem & Co environments and counters
%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{lemma}[theorem]{Lemma}
%\newtheorem{corollary}[theorem]{Corollary}
%\newtheorem{remark}[theorem]{Remark}
%\newtheorem{definition}[theorem]{Definition}
\newtheorem{equat}[theorem]{Equation}
%\newtheorem{example}[theorem]{Example}
%-------------------------------------------------------------

\begin{document}
%
\frontmatter          % for the preliminaries
%
%\pagestyle{headings}  % switches on printing of running heads
%\addtocmark{Hamiltonian Mechanics} % additional mark in the TOC
%%
%\tableofcontents
%
\mainmatter              % start of the contributions
%
\title{ Communication efficient learning from Multiple Experts}
%
%\titlerunning{Hamiltonian Mechanics}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Bojan  Kolosnjaji\inst{1} \and Huang Xiao\inst{2} 
	\and Claudia Eckert\inst{1}\inst{2} }
%Jeffrey Dean \and David Grove \and Craig Chambers \and Kim~B.~Bruce \and
%Elsa Bertino}
%
\authorrunning{Bojan Kolosnjaji et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
%\tocauthor{Ivar Ekeland, Roger Temam, Jeffrey Dean, David Grove,
%Craig Chambers, Kim B. Bruce, and Elisa Bertino}
%
\institute{Technical University of Munich, Fraunhofer AISEC\\
\email{kolosnjaji@sec.in.tum.de, huang.xiao@aisec.fraunhofer.de, eckert@sec.in.tum.de}
%\\ WWW home page:
%\texttt{https://www.aisec.fraunhofer.de/}
%\and
%Universit\'{e} de Paris-Sud,
%Laboratoire d'Analyse Num\'{e}rique, B\^{a}timent 425,\\
%F-91405 Orsay Cedex, France
}

\maketitle              % typeset the title of the contribution

\begin{abstract}
Learning from multiple sources of labeled data can increase the confidence of a machine learning system, 
if the expertise of different labelers can be assembled efficiently. However,
using a large amount of labelers in the test time can be expensive in the long run. For instance, in crowdsourcing we have to pay each labeler in order to get the annotations. In % a scenario of 
Internet-of-Things, 
obtaining data from multiple sensors for direct fusion 
incurs high communication overhead. 
%Nevertheless, in many situations a small subset of 
%%competent experts 
%data sources could already 
%%constitute a very accurate labeling system
%deliver reliable annotations,

%Therefore we need to reduce the cost of communication with other labelers effectively}. 
In this paper, we 
%design a methodolojgy
propose a method for training a classification system 
%where data labels are obtained from a small set of labelers and the labelers are weighted optimally in order to offer high classification accuracy. 
from data labeled by multiple annotators, where only a small subset of annotators are optimally chosen to reduce communication effort. We define an iterative optimization procedure where the sparsity of a client weight vector is enforced, 
%is controlled 
while reducing the overall classification error. 
%Using this procedure, 
Using a set of experiments, we show that our approach 
%obtain results close to 
can pertain the classification performance as on the complete set of labelers, while reducing 
%the number of those 
the communication effort by over 60\% on several different tasks.

%submit  abstract until 13.04

\keywords{machine learning, communication-efficiency, crowdsourcing, Internet-of-Things}
\end{abstract}
%

\section{Introduction}
Machine Learning systems depend on our ability to gather a large set of data. In a large number of scenarios, we gather labeled data from multiple separate sources with varying reliability and join this data into one machine learning model. Furthermore, many times we are limited in how many data sources we are allowed to use, because of budget constraints, limited processing power or insufficient communication bandwidth.

First important fitting scenario is crowdsourcing. When using crowdsourcing systems such as Amazon Mechanical Turk, we can gather labelled data from multiple annotators. After gathering this training data, we need to join the labelling decisions and train one machine learning system while leveraging all of the annotations. Relying on a concept of \textit{wisdom of crowds}, we can assume that if we gather data from a sufficient number of people, on average most of them will give correct labelling decision. However, gathering data from a large amount of annotators can be expensive, as workers on Amazon Mechanical Turk require payment for their work. There fore we would benefit from a procedure of prior determination of the expertise of the annotators. In this case, the joining process can be optimized by estimating the expertise of the annotators, selecting the most reliable ones and determining weights of their decisions.

Moreover, in a rising area of \textit{Internet of Things} (IoT)we can have similar problems. In IoT systems, a large number of sensors provide information about the environment. Communication among multiple sensors can be costly in terms of processing power. Furthermore, this communication can be done among sensors that are limited with power consumption constraints.

There are multiple papers that deal with estimating the expertise of labelers when using \textit{wisdom of crowds}. For example, Welinder et al. \cite{} determine groups of annotators with similar expertise, and find particular phenomena such as "schools of thought" and specialist experts in certain types of image labeling tasks. Tian and Zhu~\cite{} investigate the annotator behavior in executing more complex labeling tasks, such as answering if an image is beautiful or if it contains a car~\cite{}. Zhang et al.~\cite{} propose a method to filter out novice labelers. However, their method is specific to Gaussian Mixture Model. Rodrigues et al.~\cite{} determine a similar procedure specific to Gaussian Processes. Bi et al. argue that the factors that determine the annotators' labeling performance, apart from the expertise, can also be the worker's dedication to the task, his/her default labeling judgement and sample difficulty. However, they do not go further in determining the optimal subset of annotators. On the other hand, Li and Liu~\cite{} propose a combinatiorial procedure to find this optimal subset. However, such combinatorial procedures are very resource-intensive, especially for a large initial set of annotators.

On the other hand, there are are also multiple more theoretical frameworks. For example, Wang et al.~\cite{} propose a selection of sensors based on Directed Acyclic Graphs (DAGs) in resource-constrained scenarios. However, they do not deal with annotating data, so their ideas does not completely fit to our problem.

We design a methodology that both selects an optimal set of annotators from a set of training data and automatically determines the weights of the selected annotators. By forcing sparsity on our annotator weight vector, we get an accurate classification system that works well even with annotations from a small set of highly competent annotators.

In summary, we make the following contributions:
\begin{itemize}
\item We propose a methodology for selecting an optimal subset of workers in a classification system with multiple labelers
\item We create a procedure for selecting and determine weights of labelers independent of type machine learning methods selected for modeling the annotating process
\item Using a set of experiments with real and synthetic data, we show that we can obtain accurate classification system with a learned annotator subset.
\end{itemize}

\section{Methodology}

\section{Results}

\section{Discussion}

\section{Conclusion}

% ---- Bibliography ----
%
\bibliographystyle{plain}
\bibliography{draft} % use this
%\clearpage
%\addtocmark[2]{Author Index} % additional numbered TOC entry
%\renewcommand{\indexname}{Author Index}
%\printindex
%\clearpage
%\addtocmark[2]{Subject Index} % additional numbered TOC entry
%\markboth{Subject Index}{Subject Index}
%\renewcommand{\indexname}{Subject Index}
%\input{subjidx.ind}
\end{document}